---
title: 'thesis_survey_analysis'
author: 'Adam Hardaker'
date: '03/27/2022'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE, tidy.opts = list(width.cutoff = 60),
                      fig.height = 9, fig.width = 7)
options(scipen=10000)
```

import time!
```{r message=FALSE, warning=FALSE}
rm(list=ls())

library(lmtest)
library(sandwich)
library(Hmisc)
library(tinytex)
library(ggplot2)
library(tidyverse)
library(stargazer)
library(fastDummies)
library(gridExtra)
library(readxl)
library(writexl)
library(plotly)

theme_set(theme_gray())
setwd('C:/Users/Beemo/OneDrive/BCB/Thesis/green_identity_environmental_impact')
```


IF FIRST TIME RUNNING THIS CODE:

## load raw data

Since there are 2 rows of headers in the file, we read in the column names and 
then the body of the data, adding the names back on top

Note: data will not load if doc is currently open
```{r}
# get column names
cnames <- read_excel('data_material_footprint_2022-03-20_10-38.xlsx', n_max = 0) %>%
  names()

# read the rest, adding col_names manually
sosci_survey_raw <- read_excel('data_material_footprint_2022-03-20_10-38.xlsx',
                               skip = 2,
                               col_names = cnames)
```


## Preparing raw data for analysis

The following creates a 'derived' dataset. To do this, we:
  1) select case, language, SWB, ID, SAES, survey Qs, SD controls
  2) convert string values to integers
  ^ un-convertable values are turned into 'NA'
  3) filter for 'valid' cases, where all variables of interest have an answer
      including: SWB, identity, SAES, and req'd survey questions

```{r}
# select variables
sosci_survey_derived <- sosci_survey_raw %>%
  select(c('CASE', 9:13, 37:133, 149:156,158))

# converts strings to numerics
# also clean housing size data, by turning answers of '?' and
# 'I dont know' into NA values.
sosci_survey_derived[c(1:112)] <- lapply(sosci_survey_derived[c(1:112)], as.integer)

# filter for 'valid' cases, where all variables of interest have an answer
# including: SWB, identity, SAES, and req'd survey questions, and all sociodemogrpahic variables used in reg
sosci_survey_derived <- sosci_survey_derived %>%
  filter(sosci_survey_derived[2:7] > 0 &
           HO04 >= -9 &
           HO05 >= -9 &
           HO06 >= -9 &
           HO07 >= -9 &
           CS07 >= -9 &
           NU01 >= -9 &
           NU03 >= -9 &
           LE03 >= -9 &
           MO04 >= -9 &
           SD02_01 >= 0 &
           SD03 >= 0 &
           SD04 >= 0 &
           SD05 >= 0 &
           SD06 >= 0 &
           SD07 >= 0 &
           SD08 >= 0 &
           SD09 >= 0)

# if a respondent selected fuel type but skipped the fuel usage
# (at least one person did) then change this to the 'I don't know' answer
sosci_survey_derived['MO06'][sosci_survey_derived['MO06'] == -9] <- -1
  
# clear raw data
rm(sosci_survey_raw, cnames)

# save data. note any filepath
write_xlsx(sosci_survey_derived, 'sosci_survey_derived.xlsx')
```

We run the python script on the derived data. This had to be ran a few 
times since the operation kept timing-out. The script outputs a spreadsheet 
titled 'impact_results'.

Now, we read the python-produced spreadsheet and join it to the derived R data,
while removing most raw survey answers.

We will also transform the data into a more usable formal. To do this, we:
- create an overall identity score (reserve scoring the second 2 ID questions)
- replace codes with nominal variable categories for:
    - country
    - gender
    - ownership
    - marital status
    - employment
    - nationality
- replace survey codes with ordinal categories for:
    - education (low/med/high)
    - income
    
```{r}
impact_results <- read_excel('impact_results.xlsx')
sosci_survey_merged <- sosci_survey_derived %>%
  left_join(impact_results, by = 'CASE') %>% # merges data
  rename(identity1 = ID02_01, # renames variables
         identity2 = ID02_02,
         identity3 = ID02_03,
         identity4 = ID02_04,
         SAES = SA01,
         country = HO02,
         age = SD02_01,
         gender = SD03,
         income = SD04,
         education = SD05,
         ownership = SD06,
         employment = SD07,
         marital_status = SD08,
         health = SD09,
         postal_code = SD11_01,
         impact_housing = EI_HO,
         impact_consumer = EI_CS,
         impact_nutrition = EI_NU,
         impact_leisure = EI_LE,
         impact_mobility = EI_MO,
         impact_vacation = EI_VA,
         impact_total = EI_SUM)

# reverse score identity3 and identity4, which are on a 5pt scale
sosci_survey_merged[ ,4:5] <- 6 - sosci_survey_merged[ ,4:5]

# reverse coding health (not sure how this one slipped past :P)
sosci_survey_merged$health <- 6 - sosci_survey_merged$health

# now that we have valid cases, assume default value of '0'
#for NA values for # of children
sosci_survey_merged$HO03_02[is.na(sosci_survey_merged$HO03_02)] <- 0
sosci_survey_merged$HO03_03[is.na(sosci_survey_merged$HO03_03)] <- 0

# add dummy variables for homeowner, female, unemployed, retired, low and high educated
# and adds number of children and identity score variable
sosci_survey_merged <- sosci_survey_merged %>%
  mutate(country = case_when(country == 3 ~ 'Brazil',
                             country == 4 ~ 'China',
                             country == 5 ~ 'Denmark',
                             country == 6 ~ 'Germany',
                             country == 10 ~ 'UK',
                             country == 18 ~ 'Norway',
                             country == 25 ~ 'Spain',
                             country == 30 ~ 'USA',
                             country == 32 ~ 'other Asian country'),
         d_homeowner = case_when(ownership == 1 |
                             ownership == 2 |
                             ownership == 3 |
                             ownership == 5 ~ 0,
                             ownership == 4 ~ 1),
         ownership = case_when(ownership == 1 ~ 'rented apartment',
                             ownership == 2 ~ 'owned apartment',
                             ownership == 3 ~ 'rented house',
                             ownership == 4 ~ 'owned house',
                             ownership == 5 ~ 'other'),
         d_female = case_when(gender == 2 |
                             gender == 3 ~ 0,
                             gender == 1 ~ 1),
         gender = case_when(gender == 1 ~ 'female',
                             gender == 2 ~ 'male',
                             gender == 3 ~ 'diverse'),
         d_living_together = case_when(marital_status == 1 |
                             marital_status == 3 | 
                             marital_status == 4 | 
                             marital_status == 5 ~ 0,
                             marital_status == 2 ~ 1),         
         marital_status = case_when(marital_status == 1 ~ 'single',
                             marital_status == 2 ~ 'married',
                             marital_status == 3 ~ 'separated',
                             marital_status == 4 ~ 'divorced',
                             marital_status == 5 ~ 'widowed'),
         d_retired = case_when(employment == 1 |
                             employment == 2 |
                             employment == 3 |
                             employment == 4 ~ 0,
                             employment == 5 ~ 1),
         d_unemployed = case_when(employment == 1 |
                             employment == 2 |
                             employment == 3 |
                             employment == 5 ~ 0,
                             employment == 4 ~ 1),
         employment = case_when(employment == 1 ~ 'full time',
                             employment == 2 ~ 'part time',
                             employment == 3 ~ 'student/apprentice',
                             employment == 4 ~ 'unemployed',
                             employment == 5 ~ 'retired',
                             employment == 6 ~ 'other'),
         d_low_ed = case_when(education == 3 |
                             education == 4 |
                             education == 10 |
                             education == 11 ~ 0,
                             education == 1 |
                             education == 9 ~ 1),
         d_high_ed = case_when(education == 1 |
                             education == 3 |
                             education == 4 |
                             education == 9 |
                             education == 10 ~ 0,
                             education == 11 ~ 1),
         income = case_when(income == 1 ~ '< €1000',
                             income == 2 ~ '€1000-€2000',
                             income == 3 ~ '€2000-€3000',
                             income == 4 ~ '€3000-€5000',
                             income == 5 |
                             income == 6 ~ '> €5000'),
         education = case_when(education == 1 | # current secondary students
                             education == 9  ~ 'low', # drop outs
                             education == 3  |  # 9yr degree
                             education == 4  |  # 10yr degree
                             education == 10 ~ 'med', # 12 year degree
                             education == 11 ~ 'high'), # Uni
         health = case_when(health == 1 ~ 'very bad',
                             health == 2 ~ 'bad',
                             health == 3 ~ 'average',
                             health == 4 ~ 'good',
                             health == 5 ~ 'excellent'),
         num_children = HO03_02 + HO03_03,
         identity_score = rowMeans(sosci_survey_merged[2:5]),
         SAES = SAES - 1)

# remove unnecessary variables
sosci_survey_merged <- sosci_survey_merged %>%
  select(CASE, SAES, country, 104:128) # removes unnecessary variables

# makes NA values for income for those who didn't want to select an option
sosci_survey_merged$income[sosci_survey_merged$income == -1 | sosci_survey_merged$income == -9 ] <- NA

# orders income
sosci_survey_merged$income <- factor(sosci_survey_merged$income, levels=c('< €1000', '€1000-€2000', '€2000-€3000', '€3000-€5000', '> €5000'), ordered = TRUE)

# orders health
sosci_survey_merged$health <- factor(sosci_survey_merged$health, levels=c('very bad', 'bad', 'average', 'good', 'excellent'), ordered = TRUE)

# orders education level
sosci_survey_merged$education <- factor(sosci_survey_merged$education, levels=c('low', 'med', 'high'), ordered = TRUE)

# turns all remaining values of -9 to NA for calculation
sosci_survey_merged[sosci_survey_merged == -9 ] <- NA

rm(impact_results, sosci_survey_derived)
write_xlsx(sosci_survey_merged, 'sosci_survey_merged.xlsx')
```


AFTER FIRST TIME, load merged data directly (must have correct file directories)
```{r,  message= FALSE}
rm(list=ls())
sosci_survey_merged <- read_excel('sosci_survey_merged.xlsx')
attach(sosci_survey_merged)
```

## Descriptive Statistics

Summary stats for quantitative variables
```{r, header = FALSE, results='asis'}
stargazer(as.data.frame(sosci_survey_merged[c(2,27,19,13:18,4,26)]),
          omit.summary.stat = c('p25', 'p75'),
          title = 'Descriptive statistics',
          digits = 2,
          type = 'html',
          out = 'output/summary_stats.html')

```

Summary stats for categorical variables
```{r}
cat_vars <- sosci_survey_merged %>%
  select(c(country,
           gender,
           income,
           education,
           employment,
           marital_status,
           ownership,
           health)) %>%
  apply(2, table)

print(cat_vars)
```

Pairwise Correlation Table (Pearson)

```{r, results='asis'}
# MUST BE NUMERICAL VARIABLES
cormat <- sosci_survey_merged %>%
  select(c(SAES, identity_score, impact_total, age, d_female, d_homeowner, d_unemployed, d_retired, d_low_ed, d_high_ed, d_living_together, num_children))

cortab <- round(cor(cormat, use = 'complete.obs'), 3)
cortab[upper.tri(cortab)] <- ''
cortab <- as.data.frame(cortab)

stargazer(cortab, summary = FALSE, type = 'latex', out = 'output/cortab.html')
```

Spearman Rank Correlations

```{r}
cormat2 <- sosci_survey_merged %>%
  select(c(SAES, identity_score, impact_total, impact_housing, impact_consumer, impact_nutrition, impact_leisure, impact_mobility, impact_vacation))

cortab2 <- round(cor(cormat2, use = 'complete.obs'), 3)
cortab2[upper.tri(cortab2)] <- ''
cortab2 <- as.data.frame(cortab2)

stargazer(cortab2, summary = FALSE, type = 'latex', out = 'output/EF_SAES_cor.html')
```

Getting p-values for rank correlations

```{r}
cor.test(SAES, impact_total, method = 'spearman')
cor.test(SAES, impact_housing, method = 'spearman')
cor.test(SAES, impact_consumer, method = 'spearman')
cor.test(SAES, impact_nutrition, method = 'spearman')
cor.test(SAES, impact_leisure, method = 'spearman')
cor.test(SAES, impact_mobility, method = 'spearman')
cor.test(SAES, impact_vacation, method = 'spearman')

cor.test(identity_score, impact_total, method = 'spearman', exact = FALSE)
cor.test(identity_score, impact_housing, method = 'spearman', exact = FALSE)
cor.test(identity_score, impact_consumer, method = 'spearman', exact = FALSE)
cor.test(identity_score, impact_nutrition, method = 'spearman', exact = FALSE)
cor.test(identity_score, impact_leisure, method = 'spearman', exact = FALSE)
cor.test(identity_score, impact_mobility, method = 'spearman', exact = FALSE)
cor.test(identity_score, impact_vacation, method = 'spearman', exact = FALSE)
```


# Data Analysis

## 1: SAES and impact

Stargazer regression table 
```{r, header= FALSE, results = 'asis'}
spec1 <- lm(impact_total ~ SAES)
spec2 <- lm((impact_total ~ SAES + age + d_female + as.factor(health) + as.factor(income) + d_homeowner + d_unemployed + d_retired + d_low_ed + d_high_ed + d_living_together + num_children))

#create robust SE's
rob_se <- list(sqrt(diag(vcovHC(spec1, type = 'HC1'))),
               sqrt(diag(vcovHC(spec2, type = 'HC1'))))

stargazer(spec1, spec2,
          type='latex',
          se = rob_se,
          digits = 3,
          header = F,
          column.sep.width = '2pt',
          out = 'output/impact_SAES_reg_table.html',
          column.labels = c('Base', 'Base + Controls'),
          dep.var.caption = 'Dependent Variable: Impact',
          dep.var.labels.include = F,
         # covariate.labels = c('SAES', 'age', 'female', 'health', 'income', 'homeowner', 'unemployed', 'retired', 'low ed', 'high ed', 'num children'),
          omit.stat = c('ser', 'rsq', 'f'),
          notes = '(HC-robust standard errors in parentheses)')
```

## 2: Identity and impact

```{r}
spec3 <- lm(impact_total ~ identity_score)
spec4 <- lm((impact_total ~ identity_score + age + d_female + as.factor(health) + as.factor(income) + d_homeowner + d_unemployed + d_retired + d_low_ed + d_high_ed + d_living_together + num_children))


#create robust SE's
rob_se <- list(sqrt(diag(vcovHC(spec3, type = 'HC1'))),
               sqrt(diag(vcovHC(spec4, type = 'HC1'))))

stargazer(spec3, spec4,
          type='latex',
          se = rob_se,
          digits = 3,
          header = F,
          column.sep.width = '2pt',
          out = 'output/impact_identity_reg_table.html',
          column.labels = c('Base', 'Base + Controls'),
          dep.var.caption = 'Dependent Variable: Identity',
          dep.var.labels.include = F,
         # covariate.labels = c('identity', 'age', 'female', 'health', 'income', 'homeowner', 'unemployed', 'retired', 'low ed', 'high ed', 'num children'),
          omit.stat = c('ser', 'rsq', 'f'),
          notes = '(HC-robust standard errors in parentheses)')
```


Plot regressions

```{r, message = FALSE}
plot_SAES <- sosci_survey_merged %>%
  ggplot(aes(x=SAES, y = impact_total)) +
  geom_point() +
  geom_jitter(alpha = 0.6, height = 0.1, width = 0.05) +
  scale_x_continuous(limits = c(1,10), breaks = seq(0, 10, 1)) +
  scale_y_continuous(breaks = seq(0, 150, 10)) +
  labs(x = 'Self-assessed sustainability', y = 'Environmental impact (tons)') +
  geom_abline(intercept = 32.751, slope = -2.230, size=1, color = 'red')

plot_impact <- sosci_survey_merged %>%
  ggplot(aes(x=identity_score, y = impact_total)) +
  geom_point() +
  geom_jitter(alpha = 0.6, height = 0.1, width = 0.05) +
  scale_x_continuous(limits = c(2,5), breaks = seq(0, 5, .5)) +
  scale_y_continuous(breaks = seq(0, 150, 10)) +
  labs(x = 'Identity', y = 'Environmental impact (tons)')
 # stat_smooth(method = lm, se = FALSE, fullrange = TRUE)

grid.arrange(plot_SAES, plot_impact, ncol=2)
```

Stacked and reordered bar chart for total impact by sector

```{r}
sosci_survey_merged %>%
  select(CASE, impact_housing, impact_nutrition, impact_mobility, impact_consumer, impact_vacation, impact_leisure) %>%
  plot_ly(x = ~row.names(sosci_survey_merged), y= ~impact_vacation, type = 'bar', name= 'vacation', marker = list(color = '#f94144')) %>%
  add_trace(y = ~impact_mobility, name = 'mobility', marker = list(color = '#f3722c')) %>%
  add_trace(y = ~impact_leisure, name = 'leisure', marker = list(color = '#f9c74f')) %>%
  add_trace(y = ~impact_nutrition, name = 'nutrition', marker = list(color = '#43aa8b')) %>%
  add_trace(y = ~impact_consumer, name = 'consumer', marker = list(color = '#577590')) %>%
  add_trace(y = ~impact_housing, name = 'housing', marker = list(color = '#277da1')) %>%
  layout(yaxis = list(title = 'Environmental Impact (tons of material intensity)'),
         xaxis = list(categoryorder = 'total ascending', title = '', showticklabels=FALSE),
         legend = list(title=list(text='<b> Sector </b>')),
         bargap = 0,
         barmode = 'stack')
```

END/////////////////////////////////////////////////////////////////////////////

